### 2011-02-07 docs/architecture/architecture.tex
get the whole thing building again
more updates throughout

@@

\begin{figure}[h]
\centering
\includegraphics[width=5.0in]{\archfigpath/maestro_directory2}
\caption[MAESTRO directory structure]
{The basic MAESTRO directory structure}
\end{figure}

All the files needed to build MAESTRO are contained in the {\tt
fParallel} directory structure.  Briefly, the sub-directories
are:
\begin{itemize}
\item {\tt boxlib/}

@@

  The multigrid solver.

@@

MAESTRO models problems that are in tight hydrostatic equilibrium.
The fluid state is decomposed into a 1-d radial base state that
describes the hydrostatic structure of the star or atmosphere, and a
2- or 3-d Cartesian full state, that captures the departures from
hydrostatic equilibrium.  Two basic geometries are allowed.  A {\em plane-parallel}
geometry assumes that the domain is thin compared to the radius of curvature
of the star, and therefore the 1-d base state is perfectly aligned with
the Cartesian state.  A {\em spherical} geometry is for modeling an entire star.  Here,
the 1-d base state is not aligned with the Cartesian state.  Figure~\ref{fig:base_state}
shows this geometry.

\begin{figure}[tb]
\centering
\includegraphics[height=2.0in]{base_grid} \hspace{0.5in}
\includegraphics[height=2.0in]{base_spherical}
\caption{\label{fig:base_state} (Left) For multi-level problems
  in planar geometry, we force a direct alignment between the radial
  array cell centers and the Cartesian grid cell centers by allowing
  the radial base state spacing to change with space and time.
  (Right) For multi-level problems in spherical geometry, since there
  is no direct alignment between the radial array cell centers and
  the Cartesian grid cell centers, we choose to fix the radial base
  state spacing across levels. Figure taken from~\cite{multilevel}}
\end{figure}

MAESTRO can use adaptive mesh refinement to focus resolution on
complex regions of flow.  For Cartesian/plane-parallel geometries, all
cells at the same height must be at the same level of refinement.
This restriction is to allow for the base state to directly align with
the Cartesian state everywhere.  For spherical geometries, there is no
such restriction (again, see Figure~\ref{Fig:base_state}).

@@

@@

@@

\begin{figure}[h]
\centering
\includegraphics[width=6.5in]{\archfigpath/data_loc2}
\caption[Data-centerings on the grid]
  {\label{fig:dataloc} Some of the different data-centerings:
  (a) cell-centered, (b) nodal in the $x$-direction, and (c) nodal in
  both the $x$- and $y$-directions.  Note that for nodal data, the
  integer index corresponds to the lower boundary in that direction.
  In each of these centerings, the red point has the same indices:\ (1,2).
  Not shown is the case where data is nodal in the $y$-direction only.}
\end{figure}

@@

\begin{figure}[h]
\centering
\includegraphics[width=4.0in]{\archfigpath/index_grid2}
\caption[Single-level grid structure]
{\label{fig:boxes} Three boxes that comprise a single level.  At this
  resolution, the domain is 20$\times$18 zones.  Note that the
  indexing in BoxLib starts with $0$.}
\end{figure}

>>>>

MAESTRO models problems that are in tight hydrostatic equilibrium.
The fluid state is decomposed into a 1-d radial base state that
describes the hydrostatic structure of the star or atmosphere, and a
2- or 3-d Cartesian full state, that captures the departures from
hydrostatic equilibrium.  Two basic geometries are allowed.  A {\em
  plane-parallel} geometry assumes that the domain is thin compared to
the radius of curvature of the star, and therefore the 1-d base state
is perfectly aligned with the Cartesian state.  A {\em spherical}
geometry is for modeling an entire star.  Here, the 1-d base state is
not aligned with the Cartesian state.  Figure~\ref{fig:base_state}
shows these geometries.

\begin{figure}[tb]
\centering
\includegraphics[height=2.0in]{\archfigpath/base_grid} \hspace{0.5in}
\includegraphics[height=2.0in]{\archfigpath/base_spherical}
\caption{\label{fig:base_state} MAESTRO geometries, showing both the
  1-d base state and the full Cartesian state.  (Left) For multi-level
  problems in planar geometry, we force a direct alignment between the
  radial array cell centers and the Cartesian grid cell centers by
  allowing the radial base state spacing to change with space and
  time.  (Right) For multi-level problems in spherical geometry, since
  there is no direct alignment between the radial array cell centers
  and the Cartesian grid cell centers, we choose to fix the radial
  base state spacing across levels. Figure taken
  from~\cite{multilevel}.}
\end{figure}

MAESTRO can use adaptive mesh refinement to focus resolution on
complex regions of flow.  For Cartesian/plane-parallel geometries, all
cells at the same height must be at the same level of refinement.
This restriction is to allow for the base state to directly align with
the Cartesian state everywhere.  For spherical geometries, there is no
such restriction (again, see Figure~\ref{fig:base_state}).

@@

\begin{figure}[t]
\centering
\includegraphics[width=5.0in]{\archfigpath/maestro_directory2}
\caption[MAESTRO directory structure]
{The basic MAESTRO directory structure}
\end{figure}

All the files needed to build MAESTRO are contained in the {\tt
  fParallel} directory structure.  Problems piece together various
MAESTRO directories, choosing a reaction network, equation of state,
and conductivity routine to build an executable.  Briefly, the
sub-directories are:
\begin{itemize}
\item {\tt boxlib/}

@@

  The multigrid solver, with support for both cell-centered and
  node-centered data.

@@

@@

If the {\tt helmeos} equation of state is chosen for the problem, then
the data table for the equation of state, {\tt
  extern/EOS/helmeos/helm\_table.dat} needs to be copied to the
directory in which the code is run.

@@

\item {\tt test\_particles}

  This test exercises the particle advection routine.  A simple
  circular velocity field, with the magnitude increasing with radius
  from the center is initialized.  A number of particles are then
  initialized at various radii from the center and they are advected
  for one period.  The particle paths should be perfect circles, and
  the final particle position should overlap with the initial
  position.

@@

\begin{figure}[t]
\centering
\includegraphics[width=6.5in]{\archfigpath/data_loc2}
\caption[Data-centerings on the grid]
  {\label{fig:dataloc} Some of the different data-centerings:
  (a) cell-centered, (b) nodal in the $x$-direction, and (c) nodal in
  both the $x$- and $y$-directions.  Note that for nodal data, the
  integer index corresponds to the lower boundary in that direction.
  In each of these centerings, the red point has the same indices:\ (1,2).
  Not shown is the case where data is nodal in the $y$-direction only.}
\end{figure}

@@

\begin{figure}[t]
\centering
\includegraphics[width=4.0in]{\archfigpath/index_grid2}
\caption[Single-level grid structure]
{\label{fig:boxes} Three boxes that comprise a single level.  At this
  resolution, the domain is 20$\times$18 zones.  Note that the
  indexing in BoxLib starts with $0$.}
\end{figure}
